### 2025-01-27 test/opnorm/faster_opnorm.qmd
make opnorm test

---
title: Calculating the Operator Norm in Julia
date: 27 January 2025
format: html
engine: julia
editor:
  render-on-save: true
---

# Context
This document tests different ways to calculate the operator norm of a matrix in Julia.

The operator norm is needed to calculate the step-size used in each gradient descent update, so it is important to ensure this operation fast.

# Background

We define the operator norm of a matrix $A\in\mathbb{R}^{I\times J}$ to be
$$
\left\lVert A \right\rVert_{op} = \sup_{\lVert v \rVert_{2}} \left\lVert A v\right\rVert_{2}.
$$

It is true that
$$
\left\lVert A^\top \right\rVert_{op} = \left\lVert A \right\rVert_{op}
$$

and

$$
\left\lVert A^\top A \right\rVert_{op} = \left\lVert A A^\top  \right\rVert_{op} = \left\lVert A \right\rVert_{op}^2.
$$

# Tests

Let's look at a few ways we can find the norm of a matrix in Julia. To ensure a fair comparison, we will wrap all of these in a function, and test them out on three types of dense matrices: tall, wide, and square.

## Possible routines
```{julia}
using Pkg; Pkg.activate(".")
Pkg.add(["LinearAlgebra","Random","Arpack","BenchmarkTools"])
#Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorDecomposition.jl")
```
```{julia}
using Random: rand, randn
using LinearAlgebra: svdvals, opnorm, Symmetric
using Arpack: svds
using BlockTensorDecomposition: slicewise_dot
using BenchmarkTools
```

Now for the set of function
```{julia}
fopnorm1(X) = opnorm(X)
fopnorm2(X) = sqrt(opnorm(X'X))
fopnorm3(X) = sqrt(opnorm(X*X'))
fopnorm4(X) = sqrt(opnorm(Symmetric(X'X)))
fopnorm5(X) = sqrt(opnorm(Symmetric(X*X')))

fopnorms = [fopnorm1,fopnorm2,fopnorm3,fopnorm4,fopnorm5]

svdvals_top(X) = svdvals(X)
fsvdvals1(X) = svdvals_top(X)
fsvdvals2(X) = sqrt(svdvals_top(X'X))
fsvdvals3(X) = sqrt(svdvals_top(X*X'))
fsvdvals4(X) = sqrt(svdvals_top(Symmetric(X'X)))
fsvdvals5(X) = sqrt(svdvals_top(Symmetric(X*X')))

fopnorms = [fsvdvals1,fsvdvals2,fsvdvals3,fsvdvals4,fsvdvals5]

svds_top(X) = svds(X;nsv=1)[1].S[1]
fsvds1(X) = svds_top(X)
fsvds2(X) = sqrt(svds_top(X'X))
fsvds3(X) = sqrt(svds_top(X*X'))
fsvds4(X) = sqrt(svds_top(Symmetric(X'X)))
fsvds5(X) = sqrt(svds_top(Symmetric(X*X')))

fsvdss = [fsvds1,fsvds2,fsvds3,fsvds4,fsvds5]
```
```{julia}
b = @benchmark fopnorm1(A) setup=(A=randn(10,100))
```

```julia
dims = [
    (100, 10),
    (10, 100),
    (100, 100),
    (10, 10)
]

for (N, M) in dims
    A = randn(N, M)
    b = @benchmark opnorm(A)
    display(b)
end
```